---
title: "New method estmating R, G and Z from size distribution"
author: "Charles Hinchliffe"
output: html_notebook
editor_options: 
  chunk_output_type: console
---



This .rmd contains the new method for estimating recruitment $R$, growth $G$ and mortality $Z$, from a size distribution of larval fishes.

First we generate data based on the the following assumptions:
* $\bar{z}$ is average mortality,
* There is a linear relationship between age and size, such that $s(a) = s_0 + \bar{g}\, a,$ where $\bar{g}$ is average growth rate across the individuals lifetime. 
This implies that $\frac{\delta\, g}{\delta \,s} = 0$, and $a = (s - s_0)/\bar{g}$, leading to a predicted size distribution of 

$$n(s,t) = \frac{R(t)}{g(s_0)} \, \exp\left(- \frac{\bar{z} \, (s - s_0)}{\bar{g}} \right)$$

We then use a combination of

```{r}
library(rstan)
library(tidyverse)

#Data generation and manipulation functions
source("R/generate_sample.R")
source("R/trim_data_to_mode.R")
source("R/create_bins.R")
source("R/data_for_stan_age.R")
source("R/data_for_stan_size.R")
source("R/data_for_stan_size_growth_without_var.R")
source("R/data_for_stan_size_growth.R")

#Plotting functions
source("R/plot_age_dist.R")
source("R/plot_size_dist.R")
```


## Simulating larval fish population 
First generate a population using the simulate_catch_data() function.
```{r}
#for plotting
pars <- list(R = 1000,
             s0_av = 3,
             log10s0_sd = 0,
             g_av = 0.21,
             log10g_sd = 0,
             z_av = 0.25, 
             log10z_sd = 0,
             bin_width_age = 1,
             bin_width_size = 0.1)

#simulate population
data <- simulate_catch_data(R = 1000,
                            s0_av = 3,
                            log10s0_sd = 0,
                            g_av = 0.21,
                            log10g_sd = 0,
                            z_av = 0.25, 
                            log10z_sd = 0)

plot(size_sampled ~ age, data= data)
```

## Plotting functions
Check plot againt theoretical age distribution
```{r}
plot_age_dist(data, pars)
```


Check plot againt theoretical size distribution
```{r}
plot_size_dist(data, pars)
```

# Bin data 

Create age and size bin variable for data at smallest possible resolution (based on measurement technique)

```{r}
data_binned <-  create_bins(data,
                    bin_width_age = 1,
                    bin_width_size = 0.1)
```


## Basic Age model 

Generate data for mortality age model
```{r}
data_for_stan_age2 <- data_for_stan_age(data = data_binned, bin_width_age = 1)
```

Run stan model
```{r}
ret_age <- stan_model(file="stan/mortality_age.stan")

fit_age <- sampling(ret_age, data=data_for_stan_age2, iter = 999, chains = 3)

# stan_trace(fit_age, inc_warmup = TRUE)
# stan_hist(fit)
print(fit_age, probs=c(0.025, 0.5, 0.975))

fit_Z_age <- mean(rstan::extract(fit_age, pars="Z")[["Z"]])
fit_R_age <- mean(rstan::extract(fit_age, pars="R")[["R"]])

# plot fitted object
fitted_age <- list(
              R = fit_R_age,
              z_av = fit_Z_age,
              bin_width_age = pars$bin_width_age
)

plot_age_dist(data, pars, fitted_age)
```

## Basic Size mdoel

Estimating mortality from size (with known G).

Generate data for mortality size model
```{r}
data_for_stan_size2 <- data_for_stan_size(data = data_binned, bin_width_size = 0.1, s0_av = 3)
```

Run stan model and observe output
```{r}
ret_size <- stan_model(file="stan/mortality_size_known_g.stan")

fit_size <- sampling(ret_size, data = data_for_stan_size2, iter = 999, chains = 3)

print(fit_size, probs=c(0.025, 0.5, 0.975))
fit_Z_size <- mean(rstan::extract(fit_size, pars="Z")[["Z"]])
fit_R_size <- mean(rstan::extract(fit_size, pars="R")[["R"]])
fit_g_size <- mean(rstan::extract(fit_size, pars="g")[["g"]])

# plot fitted object
fitted_size <- list(
              R = fit_R_size,
              s0_av = pars$s0_av,
              z_av = fit_Z_size,
              g_av = fit_g_size,
              bin_width_size = pars$bin_width_size
)

plot_size_dist(data, pars, fitted_size)

```


## Size model also estimating growth

Now try estimating growth rates too (without variance). 

Produce data for mortality_size_growth model without variance in age-length
```{r}
data_for_stan_size_growth_without_var2 <- data_for_stan_size_growth_without_var(data = data_binned,
                                                  bin_width_size = 0.1,
                                                  rough_bin_width = 1.5,
                                                  n_samples_each = 3)
```

Now fit stan model
```{r}
ret_size_growth <- stan_model(file="stan/mortality_size_growth_estimatel0.stan")

fit_size_growth <- sampling(ret_size_growth, data= data_for_stan_size_growth_without_var2, iter = 999, chains = 3)

print(fit_size_growth, probs=c(0.025, 0.5, 0.975))
fit_Z_sg <- mean(rstan::extract(fit_size_growth, pars="Z")[["Z"]])
fit_R_sg <- mean(rstan::extract(fit_size_growth, pars="R")[["R"]])
fit_g_sg <- mean(rstan::extract(fit_size_growth, pars="g")[["g"]])
fit_l0_sg <- mean(rstan::extract(fit_size_growth, pars="l0")[["l0"]])

# plot fitted object
fitted_sg <- list(
              R = fit_R_sg,
              s0_av = fit_l0_sg,
              z_av = fit_Z_sg,
              g_av = fit_g_sg,
              bin_width_size = pars$bin_width_size
)

plot_size_dist(data, pars, fitted_sg)
```


Wow, pretty good, cool!!

# Size model with var age-length

Now let's add some measurement error to age-length samples

Create data with age-length error applied

```{r}
data_for_stan_size_growth <- data_for_stan_size_growth(data = data_binned,
                                                  bin_width_size = 0.1,
                                                  rough_bin_width = 1.5,
                                                  n_samples_each = 3,
                                                  age_var = 1,
                                                  size_var = 0.1)
                                    
```


Now run size model that estimates Z, G and R!
```{r}

ret_size_growth <- stan_model(file="stan/mortality_size_growth.stan")

fit_size_growth <- sampling(ret_size_growth, data=data_for_stan_size_growth, iter = 999, chains = 3)

print(fit_size_growth, probs=c(0.025, 0.5, 0.975))

stan_hist(fit_size_growth)
stan_trace(fit_size_growth, inc_warmup = TRUE)


fit_Z_sg <- mean(rstan::extract(fit_size_growth, pars="Z")[["Z"]])
fit_R_sg <- mean(rstan::extract(fit_size_growth, pars="R")[["R"]])
fit_g_sg <- mean(rstan::extract(fit_size_growth, pars="g")[["g"]])

# plot fitted object
fitted_sg <- list(
              R = fit_R_sg,
              s0_av = pars$s0_av,
              z_av = fit_Z_sg,
              g_av = fit_g_sg,
              bin_width_size = pars$bin_width_size
)

plot_size_dist(data, pars, fitted_sg)
```


## next steps

Try generating data with variation in other parameters



