---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting_Started}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = "hold"
)

library(sizedist)
```

### An introduction on how to fit size distribution models to estimate growth and mortality

`sizedist` provides functions to simulate and fit size and age-structured populations models to data. In this *Getting Started* vignette, we will cover the simplest model that is implemented in the package. `model1` assumes mortality and growth are constants. See the Theory vignette for more details on the various models in `sizedist`

To begin, let's first start by simulating some data of a hypothetical population. In order to simulate a population, we first need to set some **hyperparameters** that are used to inform the simulation. See the XXX vignette on the definitions for each hyperparameter. `sizedist` has default hyperparameters that you can access using the `default_pars()` function. Note that these hyperparameters are organised as a **list** object. Learn more about lists <here>(https://www.tutorialspoint.com/r/r_lists.htm).  
```{r hyperparams}
library(sizedist)

#Calling this function will return a list of hyperparameters
#Tip: Assign this function into an object for later use
pars <- default_pars("model1")

pars
```

We can now use these hyperparameters to simulate some data using the `simulate_population()` function.

```{r simulate}
#Tip: Assign this function to an object so you have access to the data
data <- simulate_population(pars)

data
```

We created two plotting functions to visualise the distribution of different sizes and ages of the simulated data

```{r visualise}
plot_age_dist(data, pars, binwidth = 1)
plot_size_dist(data, pars, binwidth = 0.1)
```

