---
title: "Theory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Theory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE}
library(sizedist)
```
The `sizedist` package provides functions to both simulate and fit size and age-structured populations models to data. Let $s$ be the size of an individual, $g(s)$ be its growth rate (size per unit time) and $z(s)$ be its instantaneous mortality rate, and $n(s)$ be the size density distribution of individuals. Both growth and mortality rates can vary as a function of individual size. Given specifieid relationships, the theory of size-structured populations enables us to solve $n$ from $z$ and $g$ (de Roos 1997, Muller-Landau et al. 2006).  

# Theory

Predictions are derived from the general McKendrick-von Foerster partial differential equation (PDE) for size structured population dynamics (Von Foerster 1959; Sinko and Streifer 1967): 

$$\frac{\partial n(s)}{\partial t}=\ -\frac{\partial\left[g(s) n(s)\right]}{\partial s}-z(s) n(s).$$ 

Following de Roos 1997, the PDE can be solved to give an equation for $n$, as 
$$n\left(s\right)=n\left(s_0 \right)\ \exp{\left(-\int_{0}^{t_b(s)}{\left(\frac{\delta g\left(s(t^\prime)\right)}{\delta s}+z\left(s(t^\prime)\right)\right)dt^\prime}\right)},$$
where $n\left(s_0\right)$ is the initial density of recruits, $s(a^\prime,E)$ is the size of individuals age $t^\prime$ , and $t_b(s)$ is the age of individuals size $s$. By linking the flux of individuals across the lower bound $s_0$ of the size distribution to the rate $R$ at which new individuals arrive (i.e., birth rate), we can substitute 
$$n\left(s_0\right)= \frac{R}{g(s_0)}.$$
Further, we can reorganise eq. xxx  so that the integral is expressed in terms of size, not age. Following equation 22 of de Roos (1997), we get 
$$n\left(s\right)=\frac{R}{g\left(s\right)}\ \exp{\left(-\int_{s_0}^{s}{\frac{z\left(s^\prime\right)}{g\left(s^\prime\right)}ds^\prime}\right)}.$$

Substituting in specific functions for $z(s)$ and $g(s)$, leads to a solution for $n$ (Muller-Landau et al. 2006).

Below we outline XXX models withe exact analytic solutions. The different models will apply to different kinds of organisms. In each case, we are interested in XXX outcomes:

* Individual size as a function of time: $s(t) = s_0 + \int_{0}^{t} g\left(s(t^\prime)\right) dt^\prime$
* Survival of individual to age $t$: $Z(t) =\int_{0}^{t} z\left(s(t^\prime)\right) dt^\prime$
* Size density distribution of individuals: $n(s)$
* Total number of individuals between sizes $s_1, s_2$: $N(s_1, s_2) = \int_{s_1}^{s_2} n(s) ds$ 
* Age density distribution of individuals, $n_a(a)$
* Total number of individuals between ages $a_1, a_2$: $N_a(a_1, a_2) = \int_{a_1}^{a_2} n_a(a) da$ 

These form the basis for predicted population dynamics, given specified parameters. 


## Model 1: Constant growth and mortality

The simplest model assumes growth and mortality rates are constants, independent of size. This leads to a predicted size distribution (Muller-Landau et al. 2006) as follows:

Growth rate: $g = r.$
Mortality rate: $z = d.$

Individual size as a function of time: $s(t) = s_0 + r \, t.$

Survival of individual to age $t$: 

$$Z(t) = d \,t$$

Size density distribution of individuals: 

$$n\left(s\right)=\ \frac{R\left(t\right)}{r}\ \exp{\left(-\frac{d}{r}\left(s-s_0\right)\right)}.$$

Total number of individuals between sizes $s_1, s_2$:

$$N(s_1, s_2)=\ -\frac{R}{d}\ \left(\exp{\left(-\frac{d}{r}\ \left(s_2-s_0\right)\right)}-\exp{\left(-\frac{d}{r}\ \left(s_1-s_0\right)\right)}\right).$$
Age density distribution of individuals:

$$n_a(t)= -R \ \exp{\left(-d\ t\right)}.$$
Total number of individuals between ages $a_1, a_2$: 

$$N_a{(a_1, a_2)}=\ -\frac{R}{d}\left(\exp{\left(-d\ \ a_2\right)}-\exp{\left(-d\ \ a_1\right)}\right).$$
## Model 2: 

## Model 3: 

## Model 4: 

# Fitting models to data

The size distributions outlined above predict the number of individuals within an interval of size, i.e. $N(s_1, s_2)$. This estimate can be compared to the observed counts using a Poisson probability distribution
$$Pr(N_{obs}(s_1, s_2), N(s_1, s_2)) = \textrm{Poisson}\left(N_{obs}(s_1, s_2), N(s_1, s_2)\right).$$
