---
title: "Theory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Theory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE}
library(sizedist)
```

The `sizedist` package provides functions to both simulate and fit size and age-structured populations models to data. Let $s$ be the size of an individual, $g(s)$ be its growth rate (size per unit time) and $z(s)$ be its instantaneous mortality rate, and $n(s)$ be the size density distribution of individuals. Both growth and mortality rates can vary as a function of individual size. Given specifieid relationships, the theory of size-structured populations enables us to solve $n$ from $z$ and $g$ (de Roos 1997, Muller-Landau et al. 2006).  

# Theory

Predictions are derived from the general McKendrick-von Foerster partial differential equation (PDE) for size structured population dynamics (Von Foerster 1959; Sinko and Streifer 1967): 

$$
\frac{\partial n(s)}{\partial t}=\ -\frac{\partial\left[g(s) n(s)\right]}{\partial s}-z(s) n(s).
$$

Following de Roos 1997, the PDE can be solved to give an equation for $n$, as 
$$
n\left(s\right)=n\left(s_0 \right)\ \exp{\left(-\int_{0}^{t_b(s)}{\left(\frac{\delta g\left(s(t^\prime)\right)}{\delta s}+z\left(s(t^\prime)\right)\right)dt^\prime}\right)},
$$
where $n\left(s_0\right)$ is the initial density of recruits, $s(a^\prime,E)$ is the size of individuals age $t^\prime$ , and $t_b(s)$ is the age of individuals size $s$. By linking the flux of individuals across the lower bound $s_0$ of the size distribution to the rate $R$ at which new individuals arrive (i.e., birth rate), we can substitute 
$$
n\left(s_0\right)= \frac{R}{g(s_0)}.
$$
Further, we can reorganise eq. xxx  so that the integral is expressed in terms of size, not age. Following equation 22 of de Roos (1997), we get

$$
n\left(s\right)=\frac{R}{g\left(s\right)}\ \exp{\left(-\int_{s_0}^{s}{\frac{z\left(s^\prime\right)}{g\left(s^\prime\right)}ds^\prime}\right)}.
$$

Substituting in specific functions for $z(s)$ and $g(s)$, leads to a solution for $n$ (Muller-Landau et al. 2006).

Below we outline 2 models with exact analytic solutions. The different models will apply to different kinds of organisms. In each case, we are interested in outcomes of:

* Individual size as a function of time: $s(t) = s_0 + \int_{0}^{t} g\left(s(t^\prime)\right) dt^\prime$
* Survival of individual to age $t$: $Z(t) =\int_{0}^{t} z\left(s(t^\prime)\right) dt^\prime$
* Size density distribution of individuals: $n(s)$
* Total number of individuals between sizes $s_1, s_2$: $N(s_1, s_2) = \int_{s_1}^{s_2} n(s) ds$ 
* Age density distribution of individuals, $n_a(a)$
* Total number of individuals between ages $a_1, a_2$: $N_a(a_1, a_2) = \int_{a_1}^{a_2} n_a(a) da$ 

These form the basis for predicted population dynamics, given specified parameters. 


## Model 1: Constant growth and mortality

The simplest model assumes growth and mortality rates are constants, independent of size. This leads to a predicted size distribution (Muller-Landau et al. 2006) as follows:

Growth rate: $g = r.$
Mortality rate: $z = d.$

Individual size as a function of time: 

$$s(t) = s_0 + r \, t.$$

Survival of individual to age $t$: 

$$Z(t) = d \,t$$

### `model1a`
Age density distribution of individuals:

$$
n_a(t)= -R \ \exp{\left(-d\ t\right)}.
$$

Total number of individuals between ages $a_1, a_2$: 

$$
N_a{(a_1, a_2)}=\ -\frac{R}{d}\left(\exp{\left(-d\ \ a_2\right)}-\exp{\left(-d\ \ a_1\right)}\right).
$$

### `model1b-d`

Size density distribution of individuals: 

$$
n\left(s\right)=\ \frac{R\left(t\right)}{r}\ \exp{\left(-\frac{d}{r}\left(s-s_0\right)\right)}.
$$

Total number of individuals between sizes $s_1, s_2$:

$$
N(s_1, s_2)=\ -\frac{R}{d}\ \left(\exp{\left(-\frac{d}{r}\ \left(s_2-s_0\right)\right)}-\exp{\left(-\frac{d}{r}\ \left(s_1-s_0\right)\right)}\right).
$$
### `model1e`

Size density distribution of individuals can also be expressed:

$$
n(s) = c \ exp(-b\ s)
$$
Total number of individuals between sizes $s_1, s_2$:

$$
c * (exp(-b \ s_1)/b - exp(-b \ s_2)/b)
$$


## Model 2: Growth as power function

The second model assumes $z$ and $g$ vary as power functions of size, such that 

Growth rate: $g(s) = \overline{g} \,s^\gamma$

Mortality rate: $z(s) = \overline{z} \,s^\phi$

where scaling components $\phi$ and $\gamma$ are both = 1:

### `model2a`

If we assume the assume the size spectrum is approximately stable (i.e., $\frac{∂n}{∂t} = 0$) and expand the derivative on the right-hand side of eq. 1, the slope of the normalised abundance spectrum can be expressed (i.e., the slope of $log n$ versus $log s$) as

$$
\frac{\partial\log\,n}{\partial\log\,s} = -(\frac{\overline{z}}{\overline{g}}+1+\frac{\partial\log\,\overline{g}}{\partial\log\,s}).
$$
Therefore, the size density distribution of individuals can also be expressed as 

$$
n(s) = c\,s^{-b}
$$
where $c$ is a constant and $b = (\frac{\overline{z}}{\overline{g}}+1+\frac{\partial\log\,\overline{g}}{\partial\log\,s})$

In this expression, the total number of individuals between sizes $s_1, s_2$:

$$
N(s_1, s_2)= \frac{c}{1-b}(s_2^{1-b} - s_1^{1-b})
$$
### `model2b`

To resolve $\overline{g}$ and $\overline{z}$, using `model2b`, individual size as a function of time is given as: 

$$
s(t) = s_0 + exp(\overline{g} \,t).
$$

Survival of individual to age $t$: 

$$Z(t) = \overline{z} \,t$$

Size density distribution of individuals: 

$$n\left(s\right)=\ \frac{R\,s_0^{\overline{z} /\overline{g}}}{g ̅ }\ s^{-\gamma-\overline{z} /\overline{g}}.$$

Total number of individuals between sizes $s_1, s_2$:

$$
N(s_1, s_2)=\ -\frac{R}{\overline{z}}s_0^{\overline{z} /\overline{g}} \,(s_2^{\overline{z} /\overline{g}} - s_1^{\overline{z} /\overline{g}}) 
$$


# Fitting models to data

The size distributions outlined above predict the number of individuals within an interval of size, i.e. $N(s_1, s_2)$. This estimate can be compared to the observed counts using a Poisson probability distribution
$$Pr(N_{obs}(s_1, s_2), N(s_1, s_2)) = \textrm{Poisson}\left(N_{obs}(s_1, s_2), N(s_1, s_2)\right).$$

# References

Muller‐Landau H. C., Condit R. S., Harms K. E., Marks C. O., Thomas S. C., Bunyavejchewin S., Chuyong G., Co L., Davies S., Foster R., & Gunatilleke S. (2006) Comparing tropical forest tree size distributions with the predictions of metabolic ecology and equilibrium models. Ecology Letters 9, 589–602.

von Foerster H. (1959) Some remarks on changing populations. In: Stohlman F. (ed.) The kinetics of cellular proliferation (pp. 382-407) Grune and Stratton, New York, NY.

de Roos A. M. (1997) A gentle introduction to physiologically structured population models. In: Tuljapurkar S. & Caswell H. (ed.) Structured-population models in marine, terrestrial, and freshwater systems (pp. 119-204), Springer, Boston, MA.

Sinko J. W. & Streifer W. (1967) A new model for age-size structure of a population. Ecology 48, 910–918.
